A Makefile for Developing Containerized \LaTeX\ Technical Documents

### 2021-02-19 ms.tex
fix README.md

\begin{abstract}
	We propose a Makefile for developing containerized \LaTeX\ technical documents in POSIX-oriented operating systems.
	The Makefile allows the author to execute the code that generates variables, tables and figures (results), which are then used during the \LaTeX\ compilation, to produce either the draft (fast) or full (slow) version of the document.
	We also present various utilities that aid in automating the results generation and improve the reproducibility of the document.
	We release an open source repository of a template that uses the Makefile and demonstrate its use by developing this paper.
\end{abstract}

@@

We propose a Makefile for developing containerized \LaTeX\ technical documents in POSIX-oriented operating systems.
By using portable technologies that withstood the test of time, we ensure the stability and portability of the proposed Makefile.
We write this paper using a template that consists of the Makefile and choose Docker as the container technology and the Python programming language for generating the results.

@@

The default target runs the container for the \LaTeX\ compilation while the \textit{results} target builds the image and runs the container that generates the \textit{results}.
When the \textit{author} edits a file from \textit{results code} or \textit{\LaTeX\ code}, only the required compilation is automatically triggered with Make.
Forced compilations without change could be done using the \textit{touch} POSIX utility that updates the modification time of a file.

@@

Additionally, an \textit{author} can use the POSIX utility \textit{cmp} to compare \textit{documents} (generated in a different session, operating system or underlying hardware) byte by byte, and diffoscope for identifying sources of non-determinism as follows:
\begin{lstlisting}[language=bash, style=lststyle, caption={Test draft document version reproducibility. This can also be used as a test script when pushing or pull requesting to a remote repository.}]
# Generate the first draft document version.
make
# Backup the first resulting pdf.
cp tmp/ms.pdf tmp/ms-previous.pdf
# Update the modification timestamp of main.py.
touch main.py
# Generate the second draft document version.
make
# Compare draft document versions byte by byte.
cmp tmp/ms.pdf tmp/ms-previous.pdf
# Compare draft document versions
# in a human readable form.
diffoscope tmp/ms.pdf tmp/ms-previous.pdf
\end{lstlisting}

@@

\section{Discussion}
The proposed Makefile is based on plain text file editing and can be used with any POSIX-oriented operating system that supports containers and Make.
POSIX and Make have existed for decades and have passed the test of time, while the container technology was standardized with the Open Container Initiative and there exist alternatives such as Podman that could be used as drop-in replacements for Docker.
Moreover the Makefile can be combined with any version control system, container registry provider and text editor thus preventing `app/vendor lock-in' situations.

>>>>

\begin{abstract}
	We propose a Makefile for developing containerized \LaTeX\ technical documents.
	The Makefile allows the author to execute the code that generates variables, tables and figures (results), which are then used during the \LaTeX\ compilation, to produce either the draft (fast) or full (slow) version of the document.
	We also present various utilities that aid in automating the results generation and improve the reproducibility of the document.
	We release an open source repository of a template that uses the Makefile and demonstrate its use by developing this paper.
\end{abstract}

@@

We propose a Makefile for developing containerized \LaTeX\ technical documents.
By using portable technologies that withstood the test of time, we ensure the stability and portability of the proposed Makefile.
We write this paper using a template that consists of the Makefile and choose Docker as the container technology and the Python programming language for generating the results.

@@

The default target runs the container for the \LaTeX\ compilation while the \textit{results} target builds the image and runs the container that generates the \textit{results}.
When the \textit{author} edits a file from \textit{results code} or \textit{\LaTeX\ code}, only the required compilation is automatically triggered with Make.
Forced compilations without change could be done using the \textit{touch} utility that updates the modification time of a file.

@@

Additionally, an \textit{author} can use the \textit{cmp} utility to compare \textit{documents} (generated in a different session, operating system or underlying hardware) byte by byte, and diffoscope for identifying sources of non-determinism as follows:
\begin{lstlisting}[language=bash, style=lststyle, caption={Test draft document version reproducibility. This can also be used as a test script when pushing or pull requesting to a remote repository.}]
# Generate the first draft document version.
make
# Backup the first resulting pdf.
cp tmp/ms.pdf tmp/ms-previous.pdf
# Update the modification timestamp of main.py.
touch main.py
# Generate the second draft document version.
make
# Compare draft document versions byte by byte.
cmp tmp/ms.pdf tmp/ms-previous.pdf
# Compare draft document versions
# in a human readable form.
diffoscope tmp/ms.pdf tmp/ms-previous.pdf
\end{lstlisting}

@@

\section{Discussion}
The proposed Makefile is based on plain text file editing and can be used with any operating system that supports containers and Make.
Make has existed for decades and has passed the test of time, while the container technology was standardized with the Open Container Initiative and there exist alternatives such as Podman that could be used as drop-in replacements for Docker.
Moreover the Makefile can be combined with any version control system, container registry provider and text editor thus preventing `app/vendor lock-in' situations.
